dyngen: a multi-modal simulator for spearheading new single-cell omics analyses

### 2020-06-13 manuscript.Rmd
fix supp fig ordering

**Cell-specific network inference** (CSNI) methods predict not only which transcription factors regulate which target genes, but also aim to identify how active each interaction is in each of the cells, since interactions can be turned off and on depending on the cellular state. While a few pioneering CSNI approaches have already been developed [@aibar_scenicsinglecellregulatory_2017; @kuijjer_estimatingsamplespecificregulatory_2019; @liu_personalizedcharacterizationdiseases_2016], a quantitative assessment of the performance is until now lacking. This is not surprising, as neither real nor in silico datasets of cell-specific or even cell-type-specific interactions exist that are large enough so that it can be used as a ground-truth for evaluating CSNI methods. Extracting the ground-truth dynamic network in dyngen is straightforward though, given that we can calculate how target gene expression would change without the regulator being present. We used this ground-truth to compare the performance of three CSNI methods: LIONESS [@kuijjer_estimatingsamplespecificregulatory_2019], SSN [@liu_personalizedcharacterizationdiseases_2016] and SCENIC [@aibar_scenicsinglecellregulatory_2017]. We calculated the AUROC and AUPR score for each cell individually. Computing the mean AUROC and AUPR per dataset showed that pySCENIC significantly outperforms LIONESS + Pearson, which in turn outperforms SSN*.

>>>>

**Cell-specific network inference** (CSNI) methods predict not only which transcription factors regulate which target genes, but also aim to identify how active each interaction is in each of the cells, since interactions can be turned off and on depending on the cellular state. While a few pioneering CSNI approaches have already been developed [@aibar_scenicsinglecellregulatory_2017; @kuijjer_estimatingsamplespecificregulatory_2019; @liu_personalizedcharacterizationdiseases_2016], a quantitative assessment of the performance is until now lacking. This is not surprising, as neither real nor in silico datasets of cell-specific or even cell-type-specific interactions exist that are large enough so that it can be used as a ground-truth for evaluating CSNI methods. Extracting the ground-truth dynamic network in dyngen is straightforward though, given that we can calculate how target gene expression would change without the regulator being present. We used this ground-truth to compare the performance of three CSNI methods (Figure\ \ref{fig:scgrn}): LIONESS [@kuijjer_estimatingsamplespecificregulatory_2019], SSN [@liu_personalizedcharacterizationdiseases_2016] and SCENIC [@aibar_scenicsinglecellregulatory_2017]. We calculated the AUROC and AUPR score for each cell individually. Computing the mean AUROC and AUPR per dataset showed that pySCENIC significantly outperforms LIONESS + Pearson, which in turn outperforms SSN*.
